---
title: "Kidney Cancer Survival Analysis"
runtime: shiny
output: 
  flexdashboard::flex_dashboard:
    social: menu
    source: embed
    orientation: row
    vertical_layout: scroll
    theme: yeti
    #logo: ge_monogram_primary_white_RGB.png
    #css: iids.css
---

```{r setup, include=FALSE}
library(flexdashboard)
library(OIsurv)
library(survsim)
library(broom)
library(plotly)
library(dplyr)
library(gsubfn)
library(DT)
library(SurvCorr)

impute_missing = function(data, method="average", column_exception="days_to_death"){
  column = colnames(data)
  column_data_list = NULL
  for (i in 1:length(column)){
    column_data = data[,column[[i]]]
    if(column[[i]]!=column_exception){
      if(class(column_data)=="factor"){
          if(any(is.na(column_data))==TRUE){
            column_data = as.character(column_data)
            column_data[is.na(column_data)] = "Unknown"
          }
      } else if(class(column_data)=="integer"){
        non_missing_data = column_data[!is.na(column_data)]
        if(any(is.na(column_data))==TRUE){
          if(method=="average"){
            median_value = quantile(non_missing_data, p=0.5)
            column_data[is.na(column_data)] = median_value
          } else if(method=="random"){
            random_value = sample(non_missing_data, length(column_data)-length(non_missing_data))
            column_data[is.na(column_data)] = random_value
          } else{
            print("method specified is invalid")
          }
        }
      } else if (class(column_data)=="numeric"){
        non_missing_data = column_data[!is.na(column_data)]
        if(any(is.na(column_data))==TRUE){
          if(method=="average"){
            mean_value = mean(non_missing_data)
            column_data[is.na(column_data)] = mean_value
          } else if (method=="random"){
            random_value = sample(non_missing_data, length(column_data)-length(non_missing_data))
            column_data[is.na(column_data)] = random_value
          } else{
            print("method specified is invalid")
          }
        }
      } else{
        print(paste("data type of column:", column[i], "is not supported"))
      }
    }
    column_data_list[[i]] = column_data
  }
  imputed_data = do.call(cbind.data.frame, column_data_list)
  colnames(imputed_data) = column
  return(imputed_data)
}

SplitType = function(input_data){
  if (is.data.frame(input_data)){
    df_name = deparse(substitute(input_data))
    numerical_attribute_list = c()
    categorical_attribute_list = c()
    timeseries_attribute_list = c()
    notsupport_attribute_list = c()
    all_column = colnames(input_data)
    all_datatype = lapply(input_data, class)
    for (i in 1:ncol(input_data))
      if (all_datatype[i]=="numeric" | all_datatype[i]=="integer"){
        numerical_attribute_list = c(numerical_attribute_list,all_column[i])
      } else if (all_datatype[i]=="factor" | all_datatype[i]=="character" | is.ordered(input_data[[i]])) {
        categorical_attribute_list = c(categorical_attribute_list,all_column[i])
      } else if (all_datatype[i]=="Date" | all_datatype[i]=="POSIXlt" | all_datatype[i]=="POSIXct") {
        timeseries_attribute_list = c(timeseries_attribute_list,all_column[i])
      } else {
        notsupport_attribute_list = c(notsupport_attribute_list,all_column[i])
      }
    return(list(numerical=numerical_attribute_list, categorical=categorical_attribute_list, timeseries=timeseries_attribute_list, notsupport=notsupport_attribute_list))
  }
}

data = read.csv("data/data.csv")
data_modified = data[-which(names(data) %in% c("days_to_birth","disease_code","file_uuid", "bcr","bcr_patient_barcode","bcr_patient_uuid","project_code","day_of_dcc_upload","month_of_dcc_upload","year_of_dcc_upload","patient_withdrawal","days_to_initial_pathologic_diagnosis","age_at_initial_pathologic_diagnosis","year_of_initial_pathologic_diagnosis","additional_studies","informed_consent_verified", "icd_o_3_site","follow_ups","drugs","radiations","tumor_tissue_site","day_of_form_completion","month_of_form_completion","year_of_form_completion"))]

imputed_data = impute_missing(data_modified, method="average", column_exception="days_to_death")

```

Sidebar {.sidebar}
========================================

```{r}
br()
selectInput('disease_type', label='Select a Kidney Cancer Type', choices=NULL)
disease = as.character(unique(imputed_data$histological_type))
updateSelectInput(session, 'disease_type', choices = disease)
br()
```

<strong>For Patient Features tab exploration</strong>

```{r}
selectInput('explore_feature', label='Select a Patient Feature for Exploration', choices=NULL)
features = names(imputed_data)
features = features[features!="patient_id"]
features = features[features!="histological_type"]
features = features[features!="vital_status"]
features = features[features!="days_to_death"]
updateSelectInput(session, 'explore_feature', choices = features)

```

<strong>For Patients Survival tab exploration</strong>

```{r}
selectInput('explore_survival', label='Select a Patient Feature for Exploration', choices=NULL)
observe({
  if(!is.null(input$disease_type)){
    analysis_data = imputed_data[imputed_data$histological_type==input$disease_type,]
    s_features = SplitType(analysis_data)$categorical
    s_features_new = NULL
    for(i in 1:length(s_features)){
      if(length(unique(sapply(analysis_data[,s_features[i]], as.character)))>1){
        s_features_new[[i]] = s_features[i]
      }
    }
    s_features_new = s_features_new[!is.na(s_features_new)]
    s_features_new = c("all", s_features_new)
    s_features_new = s_features_new[s_features_new!="patient_id"]
    s_features_new = s_features_new[s_features_new!="vital_status"]
    updateSelectInput(session, "explore_survival", choices = s_features_new)
  }
})

```

<strong>For Survival Model tab selection</strong>
```{r}
selectInput("model_features", "Select features you want to include to the Survival Model", choices=NULL, multiple=TRUE)
observe({
  if(!is.null(input$disease_type)){
    analysis_data = imputed_data[imputed_data$histological_type==input$disease_type, ]
    s_features_num = SplitType(analysis_data)$numerical
    s_features_cat = SplitType(analysis_data)$categorical
    s_features_cat_new = NULL
    for (i in 1:length(s_features_cat)){
      analysis_data[,s_features_cat[i]] = sapply(analysis_data[,s_features_cat[i]], as.character)
      if(length(unique(analysis_data[,s_features_cat[i]]))>1){
        s_features_cat_new[[i]] = s_features_cat[i]
      }
    }
    s_features_cat_new = s_features_cat_new[!is.na(s_features_cat_new)]
    s_features = c(s_features_num, s_features_cat_new)
    updateSelectInput(session, "input_features", choices = s_features)
  }
})

selectInput("model_type", "Select a type of survival model for analysis", choices=c("Regression", "Tree", "Forest"))
```


<strong>Patients Features </strong> 
========================================

Row
---------------------------------------

### Patient Data Description {data-padding=10}

This dataset is collected from Nationwide Children's Hospital which observe 926 patients who experience three different types of kidney cancer (Kidney Chromophobe,  Kidney Clear Cell Renal Carcinoma, Kidney Papillary Renal Cell Carcinoma). Note that none of the patients withdraw from the experiment.

Row
----------------------------------------

### Alive Patient Rate 
```{r}
renderValueBox({
  analysis_data = imputed_data[imputed_data$histological_type==input$disease_type,]
  percentAlive = round(nrow(analysis_data[analysis_data$vital_status=='Alive',])/nrow(analysis_data)*100,2)
  valueBox(paste(percentAlive,"%",sep=""), icon="fa-user", color="#339900")
})
# gauge(percentAlive, min = 0, max = 100, symbol = '%', gaugeSectors(
#   success = c(80, 100), warning = c(40, 79), danger = c(0, 39)
# ))
```

### Average Days to Death
```{r} 
renderValueBox({
  analysis_data = imputed_data[imputed_data$histological_type==input$disease_type,]
  averageDaysDeath = as.integer(quantile(na.omit(analysis_data$days_to_death), prob=0.5))
  valueBox(averageDaysDeath, icon = "fa-calendar", color="#CC0000")
})
# gauge(averageDaysDeath, min = 0, max = max(imputed_data$days_to_death, na.rm=TRUE), gaugeSectors(
#   success = c(80, 100), warning = c(40, 79), danger = c(0, 39)
# ))
```

Row
-------------------------------------------

### Patients Features - Chart {data-width=600, data-height=600}
```{r}
renderPlotly({
  if(!is.null(input$disease_type)){
    analysis_data = imputed_data[imputed_data$histological_type==input$disease_type,]
    if(!is.null(input$explore_feature)){
      if(class(analysis_data[,input$explore_feature])=="numeric"|class(analysis_data[,input$explore_feature])=="integer"){
        plot = ggplot(analysis_data, aes_string(x="vital_status", y=input$explore_feature, fill="vital_status")) + geom_boxplot() + scale_fill_brewer(palette="BrBG", guide = FALSE) + labs(x="Patients' Status", y=input$explore_feature)
        ggplotly(plot)
      } else if (class(analysis_data[,input$explore_feature])=="factor"){
        gp_data = analysis_data %>% group_by_(.dots = lapply(c("vital_status", input$explore_feature), as.symbol)) %>% summarise(freq = length(patient_id)) %>% ungroup()
        plot = ggplot(gp_data, aes_string(x="vital_status", y=input$explore_feature, fill="freq")) + geom_tile() + scale_fill_gradient(high="#CC0000", low="#FFFFFF") + labs(x="Patients' Status", y=input$explore_feature)
        ggplotly(plot)
      } else{
        return()
      }
    }
  }
})

  
# library(rCharts)
# renderChart({
#   edgelist = ddply(imputed_data, c(input$explore_feature, "vital_status"), summarise, value = length(patient_id))
#   sankey_plot = rCharts$new()
#   sankey_plot$setLib('libraries/widgets/d3_sankey')
#   sankey_plot$setTemplate(script = "libraries/widgets/d3_sankey/layouts/chart.html")
#   sankey_plot$set(data = edgelist, nodeWidth = 15, nodePadding = 10, layout = 32)
#   sankey_plot
# })

```

### Patient Features - Table {data-width=400}
```{r}
renderDataTable({
  if(!is.null(input$disease_type)){
    analysis_data = imputed_data[imputed_data$histological_type==input$disease_type,]
    if(!is.null(input$explore_feature)){
      feature = input$explore_feature
      if(class(analysis_data[,input$explore_feature])=="numeric"|class(analysis_data[,input$explore_feature])=="integer"){
        summary_table = analysis_data %>% group_by_("vital_status") %>% summarise_(.dots = fn$list(min = "min($feature)", mean = "round(mean($feature),2)", sd = "round(sd($feature),2)", max = "max($feature)"))
        summary_table = data.frame(t(summary_table))
        colnames(summary_table) = c("Alive","Dead")
        summary_table = summary_table[2:5,]
        datatable(summary_table, class = 'cell-border stripe',filter = "none", options = list(searching=FALSE, dom = 't', pageLength=10))
        #summary_table
      } else if (class(analysis_data[,input$explore_feature])=="factor"){
        analysis_data[[input$explore_feature]] = sapply(analysis_data[[input$explore_feature]], as.character)
        summary_table = as.data.frame.matrix(table(analysis_data[,input$explore_feature],analysis_data[,"vital_status"]))
        #summary_table
        datatable(summary_table, class = 'cell-border stripe',filter = "none", options = list(searching=FALSE, dom = 't', pageLength=10))
      } else{
        return()
      }
    }
  }
})

renderPrint({
  if(!is.null(input$disease_type)){
    analysis_data = imputed_data[imputed_data$histological_type==input$disease_type,]
    if(is.null(input$explore_feature)){
      feature = input$explore_feature
      if(class(analysis_data[,feature])=="numeric"|class(analysis_data[,feature])=="integer"){
        result = kruskal.test(analysis_data[[feature]] ~ analysis_data[["vital_status"]])
        if (result$p.value > 0.05){
          paste("Alive Patients and Dead Patients have NO SIGNIFICANT different", feature, "with P-value", round(result$p.value,2), "with Kruskal Wallis Testing")
        } else{
          paste("Alive Patients and Dead Patients have SIGNIFICANT different", feature, "with P-value", round(result$p.value,2), "with Kruskal Wallis Testing")
        }
      } else if (class(analysis_data[,input$explore_feature])=="factor"){
        analysis_data[[feature]] = sapply(analysis_data[[feature]], as.character)
        analysis_data[["vital_status"]] = sapply(analysis_data[["vital_status"]], as.character)
        tw_table = table(analysis_data[[feature]], analysis_data[["vital_status"]])
        result = chisq.test(tw_table, correct=TRUE)
        if(result$p.value > 0.05){
          paste("Alive Patients and Dead Patients have NO SIGNIFICANT different", feature, "with P-value", round(result$p.value,2), "with Chi-Squared Testing")
        } else{
          paste("Alive Patients and Dead Patients have SIGNIFICANT different", feature, "with P-value", round(result$p.value,2), "with Chi-Squared Testing")
        }
      } else{
        return()
      }
    }
  }
})
```


<strong>Patients Survival </strong> 
========================================

Row
----------------------------------------

### Section Description {data-padding=10}
Kaplan Meier estimate is applied to calculate probability of patients survival over time until deaths. This section provides 
1. Survival functions in general 
2. Univariate Mantel-Haenszel Test for categorical patients features for each of the Kidney disease


Row {data-height=340}
----------------------------------------
### Patient General Survival Probability Esimate {data-width=700}
```{r}
renderPlotly({
  if(!is.null(input$disease_type)){
    analysis_data = imputed_data[imputed_data$histological_type==input$disease_type,]
    if(input$explore_survival=="all"){
      analysis_data = analysis_data[which(names(analysis_data) %in% c("vital_status","days_to_death"))]
      analysis_data$vital_status = ifelse(analysis_data$vital_status=="Alive",0,1)
      survival = Surv(analysis_data$days_to_death, analysis_data$vital_status)
      survival = survfit(survival ~ 1)
      survival = tidy(survival)
      survival_curve = ggplot(survival) + geom_line(aes(x=time, y=estimate), color="#0066CC", size=1) + geom_ribbon(aes(x=time, ymin=conf.low, ymax=conf.high), alpha=.25, fill="#CC0000") + labs(x="Days to Death", y="Kaplan Meier Survival Estimate")
      ggplotly(survival_curve, width = 300, height = 300)
    } else{
      s_feature = input$explore_survival
      analysis_data = analysis_data[which(names(analysis_data) %in% c("vital_status","days_to_death",s_feature))]
      analysis_data[[s_feature]] = sapply(analysis_data[[s_feature]], as.character)
      analysis_data$vital_status = ifelse(analysis_data$vital_status=="Alive",0,1)
      survival = Surv(analysis_data$days_to_death, analysis_data$vital_status)
      formula = as.formula(paste("survival~",s_feature))
      survival = survfit(formula, data=analysis_data)
      survival = tidy(survival)
      survival_curve = ggplot(survival) + geom_line(aes(x=time, y=estimate, color=strata), size=1) + geom_ribbon(aes(x=time, ymin=conf.low, ymax=conf.high, fill=strata), alpha=.25) + labs(x="Days to Death", y="Kaplan Meier Survival Estimate")
      ggplotly(survival_curve)
    }
    
  } else(return())
})
```

### Survival Probability Estimate
```{r}
renderTable({
  if(!is.null(input$disease_type)){
    analysis_data = imputed_data[imputed_data$histological_type==input$disease_type,]
    if(input$explore_survival=="all"){
      analysis_data = analysis_data[which(names(analysis_data) %in% c("vital_status","days_to_death"))]
      analysis_data$vital_status = ifelse(analysis_data$vital_status=="Alive",0,1)
      survival = Surv(analysis_data$days_to_death, analysis_data$vital_status)
      survival = survfit(survival ~ 1)
      survival = tidy(survival)
      survival = survival[,c('time','n.risk','n.event','estimate')]
      survival
    } else{
      s_feature = input$explore_survival
      analysis_data = analysis_data[which(names(analysis_data) %in% c("vital_status","days_to_death",s_feature))]
      analysis_data[[s_feature]] = sapply(analysis_data[[s_feature]], as.character)
      analysis_data$vital_status = ifelse(analysis_data$vital_status=="Alive",0,1)
      survival = Surv(analysis_data$days_to_death, analysis_data$vital_status)
      formula = as.formula(paste("survival~",s_feature))
      survival = survfit(formula, data=analysis_data)
      survival = tidy(survival)
      survival = survival[,c('time','n.risk','n.event','estimate','strata')]
      survival
    }
  } else(return())
})
```

Row {data-height=200}
---------------------------------------
### Log Rank Test on Survival Probability 
```{r}
renderPrint({
  if(!is.null(input$disease_type)){
    analysis_data = imputed_data[imputed_data$histological_type==input$disease_type,]
    if(input$explore_survival!="all"){
      s_feature = input$explore_survival
      analysis_data = analysis_data[which(names(analysis_data) %in% c("vital_status","days_to_death",s_feature))]
      analysis_data[[s_feature]] = sapply(analysis_data[[s_feature]], as.character)
      analysis_data$vital_status = ifelse(analysis_data$vital_status=="Alive",0,1)
      survival = Surv(analysis_data$days_to_death, analysis_data$vital_status)
      formula = as.formula(paste("survival~",s_feature))
      survdiff(formula, data = analysis_data, rho = 0) 
    }
  }
})
```


<strong> Survival Model </strong>
========================================

Row
----------------------------------------

### Survival Model Description {data-padding=10}
```{r}
renderText({
  if(input$model_type=="Regression"){
    print("The result displayed below utilize cox proportional hazard model with the assumption of proportional hazard. The model utilize the function 'coxph' under 'survival' package in R.")
  } else if(input$model_type=="Tree"){
    print("")
  }
})

```

### Survival Model Result
```{r}
# model = reactive({
#   if(!is.null(input$model_features)){
#       if(input$model_type=="Regression"){
#         
#       }
#   }
# 
# })
```
